Compare similarity between profiles of significant proteins
======================================================================
    
```{r setup, echo=FALSE}
suppressPackageStartupMessages(library(Prots4Prots))

#To avoid transformation of small numbers into scientific notation causing incompatibilities with further sorting methods (googleVis by ex.)
# oldScipen <- options(scipen=500)
```


First we set up which dataset we want to analyze.
```{r getdataset}
basedir <- '/home/chernan/Workspace/testRpackage2/Prots4Prots/inst/protocols/quantitation/Label-free/'
setwd(basedir)
suppressPackageStartupMessages(source("./load_data.R"))
dataset_list <- getDatasetDpulex246daysDMSOM4pg()

threshold_pval <- 0.05

methods.to.report <- list(
    Normalization = c("vsn05"), 
    Summary = c("NonA"), 
    Significance = c("lpe"), 
    Dataset = c("dpulex2dDMSOM4", "dpulex4dDMSOM4", "dpulex6dDMSOM4")
)

```

For each triplet normalization+summarization+statistic, we first explore the similarity between profiles of significant proteins.
We first extract significant proteins, for each triplet, and summarize the information in one table.
NAs (proteins not significantly differentially expressed for a given triplet, but significant in at least one other triplet) are replaced by 0, which is of no weight while computing manhattan distances.

How similar are the significant sets of proteins ?

Compare triplets
------------------------------------------------

We can do the comparison for all available triplets.

```{r compareAll, fig.width=10, fig.height=10}
basedir <- '/home/chernan/Workspace/testRpackage2/Prots4Prots/inst/protocols/quantitation/Label-free/outputs/'
setwd(basedir)

combined_methods.to.report <- as.matrix(expand.grid(methods.to.report))
# method.label <- paste(combined_methods.to.report[1,c(1,3)], collapse='_')
# dataset.label <- 'dpulex6dDMSOM4'

temp.df <- apply(combined_methods.to.report, 1, FUN = function(methods) {
    
    method.label <- paste(c(methods["Normalization"], methods["Significance"]), 
                          collapse='_')
    dataset.label <- methods["Dataset"]
    
    prot.names <- fData(
        dataset_list[[dataset.label]]$dataset)[, c("Majority.protein.IDs", 
                                                   "Fasta.headers")]
    
    output.files <- dir(
        paste(c(basedir, '/', threshold_pval, '/', dataset.label, '/', 
                method.label, '/temp'), 
              collapse=''), 
        pattern="all_tests_mtc_Rmd_data_20.*txt", 
        full.names=TRUE)
    time.out.order <- order(file.info(output.files)[["mtime"]], decreasing = TRUE)
    pvalues.data <- read.table(output.files[time.out.order[1]],
                               stringsAsFactors=FALSE, header=TRUE,quote='"')
    
    temp.fc <- data.frame(fold.change=pvalues.data[["fold.change"]], 
                          row.names=row.names(pvalues.data))
    if(methods["Significance"]=='samr') {
        temp.fc[(pvalues.data[["significant"]] == "STABLE"),] <- NA
    }
    else {
        temp.fc[(pvalues.data[["p.values.corrected"]] > threshold_pval),] <- NA
    }
    return.val <- data.frame(temp.fc, prot.names[, ])
    
    return(return.val)
})
all.prot.names <- unique(unlist(lapply(temp.df,'[[',"Majority.protein.IDs")))
all.prot.fasta <- unique(unlist(lapply(temp.df,'[[',"Fasta.headers")))

signif.df <- data.frame(all.prot.names, all.prot.fasta)
for(index.res in 1:length(temp.df)) {
    df.res <- temp.df[[index.res]]
    signif.df <- merge(signif.df, df.res, by.x='all.prot.names', by.y="Majority.protein.IDs", all.x=TRUE, suffixes=c('',index.res))
}
df.res <- signif.df[,c("all.prot.names","all.prot.fasta",names(signif.df)[grepl(x=names(signif.df),pattern="fold.change")])]
dataset.names <- methods.to.report$Dataset
names(df.res) <- c("Majority.protein.IDs","Fasta.headers", dataset.names)
whichok.pg <- apply(df.res, 1, FUN=function(x){any(!is.na(x[3:8]))})

significant.results <- df.res[whichok.pg,]
significant.results[is.na(significant.results)] <- 0
distances <- as.matrix(dist(t(significant.results[,dataset.names]), method="euclidean"))
if(!any(is.na(distances))) {
    plot <- display_heatmap(significant.results[,dataset.names], dist.method="euclidean", title="All combinations")
    print(plot)
}

norm.colors <- rep("black", length(dataset.names))
norm.colors[grepl("C1", names(significant.results[,dataset.names]))] <- "red"
norm.colors[grepl("C2", names(significant.results[,dataset.names]))] <- "darkred"
norm.colors[grepl("M4", names(significant.results[,dataset.names]))] <- "orange"

stat.pch <- rep(0, length(dataset.names))
stat.pch[grepl("dpulex2d", names(significant.results[,dataset.names]))] <- 8 #star
stat.pch[grepl("dpulex7d", names(significant.results[,dataset.names]))] <- 15 #square

pca.obj <- prcomp(t(significant.results[,dataset.names]))
plot(pca.obj$x[,1], pca.obj$x[,2], col=norm.colors, pch=stat.pch, xlab="PC1", ylab="PC2")
abline(v=0, h=0)

```
